---
title: "Final"
author: "MO"
date: "14/07/2021"
output: html_document
---

- You will use the following code to generate your datasets. Develop your algorithm using the "edx" set. For a final test of your final algorithm, predict movie ratings in the "validation" set (the final hold-out test set) as if they were unknown. RMSE will be used to evaluate how close your predictions are to the true values in the validation set (the final hold-out test set).

- **Important:** The validation data (the final hold-out test set) should NOT be used for training, developing, or selecting your algorithm and it should ONLY be used for evaluating the RMSE of your final algorithm. The final hold-out test set should only be used at the end of your project with your final model. It may not be used to test the RMSE of multiple models during model development. You should split the edx data into separate training and test sets to design and test your algorithm.

```{r eval=FALSE, include=FALSE}
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")

# Save R file: saveRDS(x, "x.rds")
# Load R file: x<- readRDS("x.rds")

# Save Environment: save.image("Final.Rdata")
# Load Environment: load("Final.Rdata")


library(tidyverse)
library(caret)
library(data.table)

# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")

movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(movieId),
                                            title = as.character(title),
                                            genres = as.character(genres))

movielens <- left_join(ratings, movies, by = "movieId")

# Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding")
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% 
      semi_join(edx, by = "movieId") %>%
      semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)
rm(dl, ratings, movies, test_index, temp, movielens, removed)
```

 - Quiz
```{r}
# Question # 1
dim(edx)

# Question # 2
edx %>% filter(rating== 0) %>% tally()
edx %>% filter(rating== 3) %>% tally()

# Question # 3
n_distinct(edx$movieId)

# Question # 4
n_distinct(edx$userId)

# Question # 5
str_detect(edx$genres, "Drama") %>% sum()
str_detect(edx$genres, "Comedy") %>% sum()
str_detect(edx$genres, "Thriller") %>% sum()
str_detect(edx$genres, "Romance") %>% sum()

# Question # 6
edx %>% group_by(movieId) %>% count() %>% arrange(desc(n))
edx %>% filter(movieId== 296) %>% distinct(title)

# Question # 7
edx %>% group_by(rating) %>% count() %>% arrange(desc(n))

```
 
- **NOTES:** movie_avs = LSE_rating_movies
             mu = mu_least_squares
             user_avgs = LSE_rating_users
 
- Building the Algorithm
```{r}
# 1. Create the Training & Test Sets from "edx" dataset
test_index <- createDataPartition(y = edx$rating, times = 1,
                                  p = 0.2, list = FALSE)
train_set <- edx[-test_index,]
test_set <- edx[test_index,]

test_set <- test_set %>% 
     semi_join(train_set, by = "movieId") %>%
     semi_join(train_set, by = "userId")


# 2. Create the RMSE function to evaluate how well my algorithm is doing
RMSE <- function(true_ratings, predicted_ratings){
    sqrt(mean((true_ratings - predicted_ratings)^2))
}

# 3. Create a "Baseline" algorithm in which i assume all ratings are the average rating & build on to improve RMSE
mu_baseline<- mean(train_set$rating)                                                           # Average Rating
baseline_RMSE<- RMSE(test_set$rating, mu_baseline)                                             # Baseline RMSE
RMSE_results <- tibble(method = "Baseline", RMSE = baseline_RMSE)                              # Create a table with RMSE's
RMSE_results

# 4. First Algorithm using least squares method as a replacement for a linear model. (Movie Effect)
mu_least_squares <- mean(train_set$rating) 
LSE_rating_movies<- train_set %>% group_by(movieId) %>% summarize(b_i = mean(rating - mu_least_squares))
# The below line adds the avg ratings in the training set to the "b_i" values i just created in the step above after joining the newly created LSE_rating_movies to the training set. I can explore `test_set %>% left_join(LSE_rating_movies, by='movieId') %>% as_tibble()` alone to test the output.
LSE_prediction_movies<- mu_least_squares + test_set %>% left_join(LSE_rating_movies, by='movieId') %>% .$b_i
LSE_RMSE_movies <- RMSE(LSE_prediction_movies, test_set$rating)

RMSE_results <- bind_rows(RMSE_results, tibble(method="Movie Effect Model", RMSE = LSE_RMSE_movies ))
RMSE_results %>% knitr::kable()

# 5. Second Algorithm using least squares method as a replacement for a linear model. (User Effect ADDED to the Movie effect above)
LSE_rating_users <- train_set %>% left_join(LSE_rating_movies, by='movieId') %>% group_by(userId) %>% summarize(b_u = mean(rating - mu_least_squares - b_i))
LSE_prediction_users <- test_set %>% left_join(LSE_rating_movies, by='movieId') %>% left_join(LSE_rating_users, by='userId') %>% mutate(pred = mu_least_squares + b_i + b_u) %>% .$pred
LSE_RMSE_users <- RMSE(LSE_prediction_users, test_set$rating)
RMSE_results <- bind_rows(RMSE_results, tibble(method="Movie + User Effect Model", RMSE = LSE_RMSE_users ))
RMSE_results %>% knitr::kable()

# 6. Best & Worst Movies WITHOUT "Regularization" (Including number of Ratings)
Best_10<- train_set %>% dplyr::count(movieId) %>% 
     left_join(LSE_rating_movies) %>%
     left_join(movie_titles, by="movieId") %>%
     arrange(desc(b_i)) %>% 
     select(title, b_i, n) %>% 
     slice(1:10) %>% 
     knitr::kable()

Worst_10<- train_set %>% dplyr::count(movieId) %>% 
     left_join(LSE_rating_movies) %>%
     left_join(movie_titles, by="movieId") %>%
     arrange(b_i) %>% 
     select(title, b_i, n) %>% 
     slice(1:10) %>% 
     knitr::kable()

# 7. Regularization
lambdas <- seq(0, 10, 0.25)
best_lambda <- sapply(lambdas, function(l){
     mu <- mean(train_set$rating)
     b_i <- train_set %>%
          group_by(movieId) %>%
          summarize(b_i = sum(rating - mu)/(n()+l))
     b_u <- train_set %>% 
          left_join(b_i, by="movieId") %>%
          group_by(userId) %>%
          summarize(b_u = sum(rating - b_i - mu)/(n()+l))
     predicted_ratings <- 
          test_set %>% 
          left_join(b_i, by = "movieId") %>%
          left_join(b_u, by = "userId") %>%
          mutate(pred = mu + b_i + b_u) %>%
          .$pred
     return(RMSE(predicted_ratings, test_set$rating))
})

qplot(lambdas, best_lambda)
lambda <- lambdas[which.min(best_lambda)]
lambda

RMSE_results <- bind_rows(RMSE_results,
                          tibble(method="Regularized Movie + User Effect Model",  
                                     RMSE = min(best_lambda)))

RMSE_results %>% knitr::kable()

# 8. Validation Set
b_i <- validation %>% group_by(movieId) %>% summarize(b_i = sum(rating - mu_least_squares)/(n()+lambda))
b_u <- validation %>% left_join(b_i, by="movieId") %>% group_by(userId) %>% summarize(b_u = sum(rating - b_i - mu_least_squares)/(n()+lambda))
predicted_ratings <- validation %>% left_join(b_i, by = "movieId") %>% left_join(b_u, by = "userId") %>% mutate(pred = mu_least_squares + b_i + b_u) %>% .$pred
RMSE_Validation <- RMSE(predicted_ratings, validation$rating)
RMSE_results <- bind_rows(RMSE_results,
                          tibble(method="Validation",  
                                     RMSE = RMSE_Validation))

# 9. Matrix Factorization
y<- train_set %>% select(userId, movieId, rating) %>% spread(movieId, rating) %>% as.matrix()
movie_titles <- edx %>% 
    select(movieId, title) %>%
    distinct()
y <- sweep(y, 2, colMeans(y, na.rm=TRUE))
y <- sweep(y, 1, rowMeans(y, na.rm=TRUE))
colnames(y) <- with(movie_titles, title[match(colnames(y), movieId)])
```
 
 